---
title: "Project proposal"
author: "macRoni"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
```

## 1. Introduction

General Question: Understand which airlines and airports are the most recommended and the reasons for it, and investigate what drives each type of traveller’s preferences.



## 2. Data

```{r read-files}
airline <- read_csv("/cloud/project/data/airline.csv")
airport <- read_csv("/cloud/project/data/airport.csv")
lounge <- read_csv("/cloud/project/data/lounge.csv")
seat <- read_csv("/cloud/project/data/seat.csv")
glimpse(airline)
glimpse(airport)
glimpse(lounge)
glimpse(seat)
```


## 3. Data analysis plan

How does the overall ratings for airlines and airports change depending of the time of year? Do these patterns repeat every year?

To what extent do passengers in business class give better overall ratings than passengers in economy/Premium Economy?
What specific ratings differ the most/which are the same?
Does this alternate between different airlines/airplane types?

What individual rating affects the overall rating for airports/airlines the most? Does this alter depending on the individual airline/airport?

Do airlines improve their ratings with a particular reviewer over time, if so in what category are the improvements made

Do airline providers have different ratings when flying from or to different countries

Do positive reviews on an airline and positive reviews on lounge and seat have a correlation?

Is it possible to segment the reviews into “positive” and “negative” ones and replace all the positive and negative reviews with the words “good” or “bad”?

After certain connections have been established between the different variables, we aim to look into ‘how’ (and ‘whether’) the results will change if one of the variables are altered.

Which traveller type is most prevailent during different times of the year?
