---
title: "Project proposal"
author: "macRoni"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
```

## 1. Introduction
The dataset is a Skytrax User Reviews Dataset (published August 2nd, 2015) at https://github.com/quankiquanki/skytrax-reviews-dataset .  


General Theme: Which amenities impact the airline rating the most

## 2. Data

```{r read-files, message = FALSE }
airline <- read_csv("/cloud/project/data/airline.csv")
airport <- read_csv("/cloud/project/data/airport.csv")
lounge <- read_csv("/cloud/project/data/lounge.csv")
seat <- read_csv("/cloud/project/data/seat.csv")
```

```{r}
glimpse(airline)
glimpse(airport)
glimpse(lounge)
glimpse(seat)
```

## 3. Data analysis plan

We first began by exploring the overall average ratings of the different airlines. 

```{r}
airline %>%
  group_by(airline_name) %>%
  summarise( average_overall_rating = mean(overall_rating, na.rm = TRUE)) %>%
  arrange(desc(average_overall_rating))

```

Next, we visualised the distribution of the traveller types the airline attracts.
Although, this visualization doesn't take into account the 31018 missing values. 

```{r}
airline %>%
  filter(type_traveller != "NA") %>%
  ggplot( mapping = aes( y = type_traveller)) +
  geom_bar() +
  labs( title = "Distribution of the traveller types", 
        y = "Traveller types" 
        ) +
  theme_minimal()
```


How does the overall ratings for airlines and airports change depending of the time of year? Do these patterns repeat every year? Which traveller type is prev during which time of the year?
Y: overall_rating, type_traveller
X : date

To what extent do passengers in business class give better overall ratings than passengers in economy/Premium Economy?
Does this alternate between different airplane types?
Y: overall_rating
X: type_traveller , cabin_flown, aircraft   

Hypothesis: we expect business to have a higher rating and will use the correlation between overall rating and cabin flown to validate our hypothesis. 

Do airlines improve their ratings with a particular reviewer over time, if so in which category are the improvements made
Y: overall_rating, other variables TBD after the most significant contributor is found 
X: author, date, airline_name 

Do positive reviews on an airline and positive reviews on lounge and seat have a correlation?
Y: a graph showing correlation 
X: overall_rating (in seat and lounge data set )

Hypothesis: we expect seat rating to have a stronger correlation than the lounge rating because not everyone uses the lounge. 

***Question to the instructor: Is it possible to segment the reviews into “positive” and “negative” ones (using text analysis on the content column) and replace all the positive and negative reviews with the words “good” or “bad”?***
