---
title: "Project proposal"
author: "macRoni"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(skimr)
library(tidytext)
```

## 1. Introduction
The dataset is a Skytrax User Reviews Dataset (published August 2nd, 2015) at https://github.com/quankiquanki/skytrax-reviews-dataset .  


General Theme: How do the amenities impact the flight's overall rating? 

## 2. Data

```{r read-files, message = FALSE }
airline <- read_csv("/cloud/project/data/airline.csv")
lounge <- read_csv("/cloud/project/data/lounge.csv")
seat <- read_csv("/cloud/project/data/seat.csv")
```

```{r}
glimpse(airline)
glimpse(lounge)
glimpse(seat)
```

## 3. Data analysis plan

We first began by exploring the overall average ratings of the different airlines. 

```{r}
airline %>%
  group_by(airline_name) %>%
  summarise( average_overall_rating = mean(overall_rating, na.rm = TRUE)) %>%
  arrange(desc(average_overall_rating))

```


1. How does the overall ratings for airlines change depending of the time of year? Do these patterns repeat every year? Which traveler type is prevalent during which time of the year?

First, all NAs for the response variable and the predictor were removed. 

```{r mean-ratings-airlines, message = FALSE}
airline_nonas <- airline %>%
  drop_na(overall_rating) #removing any NAs
```

Next, we want to investigate how time affects ratings. As the research question focuses on how airlines get high ratings, the five airlines with the highest rating were found. A limit of at least 100 survey submissions was imposed to increase the number of data points.

```{r five-highest-rated-airlines, message = FALSE, echo = FALSE}
airline_nonas %>%
  group_by(airline_name) %>%
  summarise(mean_overall_rating = mean(overall_rating, na.rm = TRUE),
            n = n()) %>%
  filter(n >= 100) %>%
  arrange(desc(mean_overall_rating))
```

```{r number of observations vs date, message = FALSE, echo = FALSE}
airline_nonas %>%
  filter(date >= "2010-01-01") %>%
  ggplot(aes(x = date)) +
  geom_histogram(binwidth = 60) +
  geom_hline(yintercept=1500, color = "orange") +
  labs(
    title = "Number of Observations vs Date",
    x = "Date",
    y = "Count"
  ) +
  theme_minimal()
```

These airlines were plotted against time to identify any correlations.

```{r five-highest-ratings-plot, message = FALSE, echo = FALSE}
airline_nonas %>%
  group_by(airline_name, date) %>%
  summarise(mean_overall_rating = mean(overall_rating, na.rm = TRUE)) %>%
  filter(airline_name %in% c("asiana-airlines", "garuda-indonesia", "air-astana", "bangkok-airways", "indigo-airlines"),
         date >= "2014-01-01") %>%
  ggplot(aes(x = date, y = mean_overall_rating, color = airline_name)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Overall ratings of five highest ranked airlines",
    subtitle = "From 2014 to 2015",
    x = "Date",
    y = "Overall Rating"
  ) +
  guides(color = guide_legend(title = "Airline Name")) 
  
```

As the visualization shows, all of the airlines receive ratings in between 5-10, with air-astana showing an increase at the end of 2015. For indigo-airlines, garuda-indonesia and bangkok-airlines, the ratings deteriorated at the end of 2015. Garuda-indonesia ratings stay constant throughout both years.

There is no clear pattern in each year, which makes sense as the airline companies will normally try to provide the same service throughout the year. And, this may also mean that the airlines probably don't take the comments given by 
the passengers too seriously.

1.5. Which amenity significantly contributes to the top 5 airlines' overall rating?

One can look at the individual ratings (Seat Comfort, Cabin Staff, Food Beverages, Inflight Entertainment and Money Value) and see if there is one rating that sticks out depending on the airline. 

```{r all-airlines-ratings-breakdown, message = FALSE, echo = FALSE}
airline_nonas %>%
  group_by(airline_name, date) %>%
  summarise(mean_seat_comfort_rating = mean(seat_comfort_rating, na.rm = TRUE),
            mean_cabin_staff_rating = mean(cabin_staff_rating, na.rm = TRUE),
            mean_food_beverages_rating = mean(food_beverages_rating, na.rm = TRUE),
            mean_inflight_entertainment_rating = mean(inflight_entertainment_rating, na.rm = TRUE),
            mean_value_money_rating = mean(value_money_rating, na.rm = TRUE),
            ) %>%
  filter(airline_name %in% c("asiana-airlines", "garuda-indonesia", "air-astana", "bangkok-airways", "indigo-airlines"),
         date >= "2014-01-01") %>%
  ggplot(aes(x = date)) +
  geom_smooth(aes(y = mean_seat_comfort_rating, color = "red"), se = FALSE) +
  geom_smooth(aes(y = mean_cabin_staff_rating, color = "green"), se = FALSE) +
  geom_smooth(aes(y = mean_food_beverages_rating, color = "blue"), se = FALSE) +
  geom_smooth(aes(y = mean_inflight_entertainment_rating, color = "yellow"), se = FALSE) +
  geom_smooth(aes(y = mean_value_money_rating, color = "black"), se = FALSE) +
  labs(
    title = "Seat Comfort, Cabin Staff, Food Beverages, Inflight Entertainment and 
Money Value Ratings for Top Five Airlines",
    subtitle = "From 2014 to 2015",
    x = "Date",
    y = "Rating"
  ) +
  scale_color_identity(name = "Type of Rating (Mean)",
                          breaks = c("red", "green", "blue", "yellow", "black"),
                          labels = c("Seat Comfort", "Cabin Staff", "Food Beverages", "Inflight Entertainment", "Money Value"),
                          guide = "legend") +
  facet_wrap(~airline_name, ncol = 3)
  
```
As the visualisation shows, to increase in overall rating, all individaul ratings have to increase. This can be seen in the example of Air Astana. For airlines that decreased in overall rating (bangkok-airways and indigo-airlines), only specific individual ratings decreased or had an overall worse score (especially inflight entertainment). 

# Question 2 

2. To what extent do passengers in first class give better ratings than passengers in other classes?
Also, how satisfied are the first class passengers with the services they mentioned in their comments the most about?


Let's look at the average overall rating the passengers gave for each of the classes. 
Hypothesis: We expect first class to have a higher rating since, as per the norm, first class is considered to be the most preferable one.

```{r}
airline %>%
  filter(!is.na(overall_rating), !is.na(cabin_flown) ) %>%
  group_by(cabin_flown) %>%
  summarise(mean_overall_rating = mean(overall_rating, na.rm = TRUE)) %>%
  arrange(desc(mean_overall_rating))
```

Business class has received the highest amount of rating. It is closely followed by 
First class.  

```{r}
airline %>%
  filter(!is.na(overall_rating), !is.na(cabin_flown) ) %>%
  group_by(cabin_flown) %>%
  count(cabin_flown) %>%
  arrange(desc(n))
```


Let's look into what type of travelers opt for first class cabin which may be able to explain the above. 


```{r}
airline %>%
  filter (!is.na(cabin_flown), !is.na(type_traveller), !is.na(overall_rating) ) %>%
  ggplot(aes(x = cabin_flown, fill = type_traveller)) +
  geom_bar(position ="fill") +
  coord_flip()

```

This reveals the type of travelers in each of the cabins.

For first class, as expected, the solo leisure traveller_type prevails. This adds up since
the tickets may cost significantly more than any of the other classes so it may 
not be pocket-friendly to travel in larger groups than 1.   
The distribution of travelers also reveals why first class has the smallest number of passengers since
'couple leisure' and 'family leisure' categories don't form a large section of first class cabin passengers.  


Let's look into which amenities did the first class cabin passengers mention the most about in their comments and how did they rate them? 
 
Filtered for NAs for variables that I am entirely sure of looking into.

```{r}
airline_amen_nonas <- airline %>%
  drop_na(overall_rating, value_money_rating, cabin_flown)  

```

Text analysis:

```{r}

two_words <- airline_amen_nonas %>%
  filter( cabin_flown == "First Class" ) %>%
  unnest_tokens(word, content, token = "ngrams", n = 2) %>%
  count(word, sort = TRUE) %>% 
  slice_max(n, n = 65)

#single_word <- airline_amen_nonas %>%
 # filter( cabin_flown == "First Class" ) %>%
 # unnest_tokens(word, content) %>%
 # anti_join(get_stopwords(), by = "word") %>%
 # count(word, sort = TRUE) %>% 
 # slice_max(n, n = 20)

```

The results reveal phrases like "wifi connectivity", "flight entertainment"  don't appear commonly, so I will choose to not look  into the amenity variable associated with those words as it is not ***highly*** reviewed by the first class passengers.
To be thorough, I have unnested the sentences into two tokens instead of just words. 

So, after narrowing my search using text analysis, I will filter for seat, cabin staff and food and, explore the ratings given for these. 
The amen_filtered has been extended accordingly from airline_amen_nonas. 

```{r}

amen_filtered <- airline %>%
  drop_na(overall_rating, cabin_flown, seat_comfort_rating,
          cabin_staff_rating, food_beverages_rating)  

```

```{r}
amen_filtered %>%
  filter( cabin_flown == "First Class" ) %>%
  select(overall_rating, value_money_rating, seat_comfort_rating,
          cabin_staff_rating, food_beverages_rating) %>%
  summarise(mean_seat_comfort_rating = mean(seat_comfort_rating),
            mean_cabin_staff_rating = mean(cabin_staff_rating),
            mean_food_beverages_rating = mean(food_beverages_rating),
            mean_overall_rating = mean(overall_rating))
```

Perhaps, this can mean that the mention of these services in the comments made has been fairly positive according to
the mean results yielded. After looking at variability, we can add that passengers had strong opinions about the food and beverages in the first class cabin of the flight. The mean values show that on average, the passengers were happier with the service offered by the crew instead of the seating comfort level. This is supported by the density plot as there are fewer observations < 3 for  cabin staff rating than seat_comfort rating. 


But, let's see the variability to reaffirm our claim. 

```{r}
 y <- amen_filtered %>%
  group_by(airline_name) %>%
  filter( cabin_flown == "First Class" ) %>%
  select(overall_rating, value_money_rating, seat_comfort_rating,
          cabin_staff_rating, food_beverages_rating) %>%
  summarise(mean_seat_comfort_rating = mean(seat_comfort_rating),
            mean_cabin_staff_rating = mean(cabin_staff_rating),
            mean_food_beverages_rating = mean(food_beverages_rating),
            mean_value_money_rating = mean(value_money_rating),
            mean_overall_rating = mean(overall_rating))
```

```{r}

seat_var <- y %>%
ggplot( mapping = aes(x = mean_seat_comfort_rating)) +
  geom_density() 

cabin_var <- y %>%
ggplot( mapping = aes(x = mean_cabin_staff_rating)) +
  geom_density() 

food_var <- y %>%
ggplot( mapping = aes(x = mean_food_beverages_rating)) +
  geom_density()

overall_var <-y %>%
  ggplot( mapping = aes(x = mean_overall_rating)) +
  geom_density()

```

This runs in line with the average statistics visualized in the tibble before. 

# End of question 2

Majority of the travelers have rated the airlines a 10 (5861 to be exact) but ratings 
like 8, 9, 10 are also common. Surprisingly, the extreme value of rating = 1 is also 
frequent. 


Do airlines improve their ratings with a particular reviewer over time, if so in which category are the improvements made
Y: overall_rating, other variables TBD after the most significant contributor is found 
X: author, date, airline_name 

Do positive reviews on an airline and positive reviews on lounge and seat have a correlation?
Y: a graph showing correlation 
X: overall_rating (in seat and lounge data set )

Hypothesis: we expect both the seat rating and the lounge rating to both have a positive correlation to the airline rating because both are provided by the airline so this would affect the authors overall rating of the airline.

To answer this question, only the authors that have rated the airline, lounge and seat will be selected to see if rating either the lounge or seat positively will result in the airline being rated positively. 

To start only the relevant columns are chosen, then the data sets are joined into one and any NAs are removed so that only the authors wanted remain.

```{r - selecting-relevant-columns}
airline_rating <- airline %>%
select(airline_name, overall_rating, author, date) %>%
  rename(airline_rating = "overall_rating",
         date_airline = "date")

lounge_rating <- lounge %>%
select(airline_name, overall_rating, author, date) %>%
  rename(lounge_rating = "overall_rating",
         date_lounge = "date")

seat_rating <- seat %>%
select(airline_name, overall_rating, author, date) %>%
  rename(seat_rating = "overall_rating",
         date_seat = "date")
```

```{r - joining-lounge-and-airline}
airline_lounge <- full_join(lounge_rating, airline_rating) %>%
  na.omit()

airline_lounge
```


```{r joining-seat-to-lounge-and-airline}
airline_lounge_seat <- full_join(airline_lounge, seat_rating) %>%
  na.omit()

airline_lounge_seat <- airline_lounge_seat[, c(1, 3, 5, 2, 7, 4, 6, 8)]

airline_lounge_seat
```

Next I created an overall rating column and a column that tells from which data set that rating came from, and created a new date column so that the dates from the corresponding rating is in that column. Then any duplicate ratings that occured whilst joining have been taken out.

```{r creating-overall-column}
als_overall_rating <- airline_lounge_seat %>%
  pivot_longer(cols = c("airline_rating", "seat_rating", "lounge_rating"), values_to = ("overall_rating")) %>%
   rename(rating_type = "name") %>%
  mutate(
    date = case_when(
      rating_type == "airline_rating" ~ date_airline,
      rating_type == "seat_rating"    ~ date_seat,
      rating_type == "lounge_rating"  ~ date_lounge
    )) %>%
  select(airline_name, author, rating_type, overall_rating, date) %>%
  distinct()

als_overall_rating
  
```

Next the mean is calculated for each type of rating for each author as some authors have provided multiple ratings for each type. Then any duplicates were removed.

```{r calculating-mean-where-authors-have-multiple-ratings-for-one-type}
als_with_mean <- als_overall_rating %>%
  group_by(author, rating_type) %>%
  mutate(mean_overall_rating = mean(overall_rating)) %>%
  select(airline_name, author, rating_type, mean_overall_rating) %>%
  distinct()

als_with_mean
```

Next the difference between the airline ratings and louge ratings, and the difference between the airline and seat ratings were calculated as a percentage of the the maximum rating that could have been given (10) and the absolute value was taken from each percentage to avoid negative percentages.

```{r calculating-the-difference-between-the-means-of-different-rating-types}
 als_with_mean_difference <- als_with_mean %>%
pivot_wider(names_from = rating_type, values_from = mean_overall_rating) %>%
mutate(percentage_difference_in_airline_and_lounge_rating = (airline_rating-lounge_rating)/10*100,
       percentage_difference_in_airline_and_seat_rating = (airline_rating-seat_rating)/10*100)

als_with_mean_difference
```

```{r changing-values-to-absolute}
abs_data <- als_with_mean_difference %>%
  mutate(abs_percentage_difference_in_airline_and_seat_rating = abs(percentage_difference_in_airline_and_seat_rating)) %>%
  select(airline_name, author, airline_rating, seat_rating, lounge_rating, percentage_difference_in_airline_and_lounge_rating, abs_percentage_difference_in_airline_and_seat_rating)

abs_data
```

To visualize the results a graph of each of the differences between ratings was created to try and see if there was a correlation.

```{r graph-airline-and-lounge-ratings}
abs_data %>%
  ggplot(mapping = aes(y = author, 
                      x = percentage_difference_in_airline_and_lounge_rating )) + 
  geom_bar(stat = "identity") +
  labs(title = "Difference in the Ratings between Airline and Lounge",
       y = "Author",
       x = "% Difference") 
```

```{r graph-airline-and-seat-ratings}
abs_data %>%
  ggplot(mapping = aes(y = author, 
                      x = abs_percentage_difference_in_airline_and_seat_rating )) + 
  geom_bar(stat = "identity") +
  labs(title = "Percentage Difference in the Ratings between Airline and Seat",
       y = "Author",
       x = "% Difference")
```
From the graphs we can see that the majority of the percentage differences between ratings is quite low, to look closer into detail a summary table can be produced to find the mean percentage difference which will show how much the ratings differed  

```{r}
abs_data %>%
  select(percentage_difference_in_airline_and_lounge_rating, abs_percentage_difference_in_airline_and_seat_rating, author) %>%
summary()

```

Hypothesis: we expect both the seat rating and the lounge rating to both individualy have a positive correlation to the airline rating because both are provided by the airline so this would affect the authors overall rating of the airline.


